# Motivating including surface parameters

We will fit the model on the following data:

```{r}
library(BradleyTerry2)

grass_data <- read.csv("data/matches_grass.csv")
hard_data <- read.csv("data/matches_hard.csv")
clay_data <- read.csv("data/matches_clay.csv")
grass_data
```


```{r}
surface_names <- c("grass", "hard", "clay")
surface_data <- list(grass=grass_data, hard=hard_data, clay=clay_data)
```

Fit the model

```
```{r}
print(setdiff(unique(grass_data$player1), unique(grass_data$player2)))
print(setdiff(unique(grass_data$player2), unique(grass_data$player1)))
btm_grass <- BTm(cbind(win1, win2), as.factor(player1), as.factor(player2), data=grass_data)
s_grass <- summary(btm_grass)
s_grass
```

```{r}
print(setdiff(unique(hard_data$player1), unique(hard_data$player2)))
print(setdiff(unique(hard_data$player2), unique(hard_data$player1)))
btm_hard <- BTm(cbind(win1, win2), as.factor(player1), as.factor(player2), data=hard_data)
s_hard <- summary(btm_hard)
s_hard
```

```{r}
print(setdiff(unique(clay_data$player1), unique(clay_data$player2)))
print(setdiff(unique(clay_data$player2), unique(clay_data$player1)))
btm_clay <- BTm(cbind(win1, win2), as.factor(player1), as.factor(player2), data=clay_data)
s_clay <- summary(btm_clay)
s_clay
```

Find top 10 players on each surface

```{r}
df_coeff <- as.data.frame(s_grass$coefficients)
rownames(df_coeff) <- lapply(rownames(df_coeff),
                             function(x) substr(x, 3, nchar(x)))
# sort based on coefficient
df_coeff <- df_coeff[order(df_coeff$Estimate, decreasing=TRUE),]
df_coeff[1:10,]
```

```{r}
df_coeff <- as.data.frame(s_hard$coefficients)
rownames(df_coeff) <- lapply(rownames(df_coeff),
                             function(x) substr(x, 3, nchar(x)))
# sort based on coefficient
df_coeff <- df_coeff[order(df_coeff$Estimate, decreasing=TRUE),]
df_coeff[1:10,]
```

```{r}
df_coeff <- as.data.frame(s_clay$coefficients)
rownames(df_coeff) <- lapply(rownames(df_coeff),
                             function(x) substr(x, 3, nchar(x)))
# sort based on coefficient
df_coeff <- df_coeff[order(df_coeff$Estimate, decreasing=TRUE),]
df_coeff[1:10,]
```

```{r}
sum(grass_data$win1)+sum(grass_data$win2)
sum(hard_data$win1)+sum(hard_data$win2)
sum(clay_data$win1)+sum(clay_data$win2)
```

```{r}
BTabilities(btm_grass)
```