---
output: pdf_document
---

# Surface

Use weighting by surface from Sipko and Knottenbelt.

```{r}
library(TennisBT)
library(tidyverse)

data <- TennisTidyr(2013, 2022, 2023, 20)
train <- data$train
test <- data$test
players <- data$main_names
```

They use the correlation between surfaces $a$ and $b$, according to the formula:

$$
\rho_{a,b} = \frac{\sum_{i=1}^n(a_i-\bar{a})(b_i-\bar{b})}{(n-1)s_as_b}
$$

$a_i$ = percentage of matches won by player $i$ on surface $a$ $b_i$ = percentage of matches won by player $i$ on surface $b$ $s_a$ = standard deviation of percentage of matches won on surface $a$ $s_b$ = standard deviation of percentage of matches won on surface $b$ $\bar{a}$ = mean percentage of matches won on surface $a$ $\bar{b}$ = mean percentage of matches won on surface $b$ $n$ = number of players

```{r}
n <- length(players)
win_counts <- as.data.frame(matrix(0, nrow = n, ncol = 3))
colnames(win_counts) <- c("Grass", "Clay", "Hard")
rownames(win_counts) <- players

for (player in players) {
  train_player <- train[train$winner_name == player,]
  player_count <- nrow(train_player)
  
  win_counts[player,"Grass"] <- nrow(train_player[train_player$surface == 'Grass',])/player_count
  win_counts[player,"Clay"] <- nrow(train_player[train_player$surface == 'Clay',])/player_count
  win_counts[player,"Hard"] <- nrow(train_player[train_player$surface == 'Hard',])/player_count
}
win_counts <- na.omit(win_counts)
row_sub = apply(win_counts, 1, function(row) all(row !=0 ))
win_counts <- win_counts[row_sub,]

grass_mean <- mean(win_counts$Grass)
grass_sd <- sd(win_counts$Grass)
clay_mean <- mean(win_counts$Clay)
clay_sd <- sd(win_counts$Clay)
hard_mean <- mean(win_counts$Hard)
hard_sd <- sd(win_counts$Hard)

```

```{r}
grass_clay <- sum((win_counts$Grass-grass_mean)*(win_counts$Clay-clay_mean))/((n-1)*grass_sd*clay_sd)
hard_clay <- sum((win_counts$Hard-hard_mean)*(win_counts$Clay-clay_mean))/((n-1)*hard_sd*clay_sd)
grass_hard <- sum((win_counts$Hard-hard_mean)*(win_counts$Grass-grass_mean))/((n-1)*hard_sd*grass_sd)
```

Now create a dataset with the weightings

```{r}
weighting <- as.data.frame(diag(3))
rownames(weighting) <- c("Grass", "Hard", "Clay")
colnames(weighting) <- c("Grass", "Hard", "Clay")
weighting['Grass', 'Hard'] <- grass_hard
weighting['Hard', 'Grass'] <- grass_hard
weighting['Clay', 'Hard'] <- hard_clay
weighting['Hard', 'Clay'] <- hard_clay
weighting['Clay', 'Grass'] <- grass_clay
weighting['Grass', 'Clay'] <- grass_clay
```

Now create three new columns in the data, weighting according to the match that the surface was played on.

```{r}
# Find the surface of each match in train, and add the corresponding weighting for each surface
for (i in 1:nrow(train)) {
  train[i, 'grass_weight'] <- weighting[train[i, 'surface'], 'Grass']
  train[i, 'clay_weight'] <- weighting[train[i, 'surface'], 'Clay']
  train[i, 'hard_weight'] <- weighting[train[i, 'surface'], 'Hard']
}
```

Train three different models weighted by each of the surfaces.

```{r}
library(BradleyTerry2)
btm_grass <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(grass_weight), data = train)
btm_clay <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(clay_weight), data = train)
btm_hard <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(hard_weight), data = train)

```

Finally predict the models.

```{r}
# Create a dataframe with the Bradley-Terry estimates for each surface
btm_grass_df <- as.data.frame(BTabilities(btm_grass))
btm_clay_df <- as.data.frame(BTabilities(btm_clay))
btm_hard_df <- as.data.frame(BTabilities(btm_hard))

coeff_frames <- list('Grass' = btm_grass_df, 'Clay' = btm_clay_df,
                     'Hard' = btm_hard_df)

draws <- test[, c("winner_name", "loser_name", "surface")]
draws$pred <- NA

# predict the test set using a different model based on the surface
for (i in 1:nrow(draws)) {
  player1 <- draws$winner_name[i]
  player2 <- draws$loser_name[i]
  draw_surface <- draws$surface[i]

  pred <- predict(player1, player2, as.data.frame(coeff_frames[draw_surface]))
  draws$pred[i] <- pred[2]
}
```

```{r}
score_predictions(draws)
```

These results are not as good as the time weighting model. We conduct a grid search below to find the optimal parameters. WARNING this code will take approximately 4 hours to run (machine dependent), hence it is stored in a function so that the markdown will knit.

```{r}
  grid_search_params <- function() {
  best_ijk <- c(0,0,0)
  best_accuracy <- 0
  
  for (i_ in seq(0,1,0.1)) {
    for (j_ in seq(0,1,0.1)) {
      for (k_ in seq(0,1,0.1)) {
        grass_clay <- i_
        hard_clay <- j_
        grass_hard <- k_
        
        weighting <- as.data.frame(diag(3))
        rownames(weighting) <- c("Grass", "Hard", "Clay")
        colnames(weighting) <- c("Grass", "Hard", "Clay")
        weighting['Grass', 'Hard'] <- grass_hard
        weighting['Hard', 'Grass'] <- grass_hard
        weighting['Clay', 'Hard'] <- hard_clay
        weighting['Hard', 'Clay'] <- hard_clay
        weighting['Clay', 'Grass'] <- grass_clay
        weighting['Grass', 'Clay'] <- grass_clay
        
        for (i in 1:nrow(train)) {
          train[i, 'grass_weight'] <- weighting[train[i, 'surface'], 'Grass']
          train[i, 'clay_weight'] <- weighting[train[i, 'surface'], 'Clay']
          train[i, 'hard_weight'] <- weighting[train[i, 'surface'], 'Hard']
        }
        
        btm_grass <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(grass_weight), data = train)
        btm_clay <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(clay_weight), data = train)
        btm_hard <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(hard_weight), data = train)
        
        # Create a dataframe with the Bradley-Terry estimates for each surface
        btm_grass_df <- as.data.frame(BTabilities(btm_grass))
        btm_clay_df <- as.data.frame(BTabilities(btm_clay))
        btm_hard_df <- as.data.frame(BTabilities(btm_hard))
        
        coeff_frames <- list('Grass' = btm_grass_df, 'Clay' = btm_clay_df,
                             'Hard' = btm_hard_df)
        
        draws <- test[, c("winner_name", "loser_name", "surface")]
        draws$pred <- NA
        
        # predict the test set using a different model based on the surface
        for (i in 1:nrow(draws)) {
          player1 <- draws$winner_name[i]
          player2 <- draws$loser_name[i]
          draw_surface <- draws$surface[i]
        
          pred <- predict(player1, player2, as.data.frame(coeff_frames[draw_surface]))
          draws$pred[i] <- pred[2]
        }
        
        acc <- score_predictions(draws)$accuracy
        if (acc > best_accuracy) {
          best_accuracy <- acc
          best_ijk <- c(i_,j_,k_)
          print(best_accuracy)
          print(best_ijk)
        }
      }
    }
  }
}
```

```{r}
final_grass_clay <- 0.4
final_hard_clay <- 0.8
final_grass_hard <- 0.5
```

```{r}
weighting <- as.data.frame(diag(3))
rownames(weighting) <- c("Grass", "Hard", "Clay")
colnames(weighting) <- c("Grass", "Hard", "Clay")
weighting['Grass', 'Hard'] <- final_grass_hard
weighting['Hard', 'Grass'] <- final_grass_hard
weighting['Clay', 'Hard'] <- final_hard_clay
weighting['Hard', 'Clay'] <- final_hard_clay
weighting['Clay', 'Grass'] <- final_grass_clay
weighting['Grass', 'Clay'] <- final_grass_clay

for (i in 1:nrow(train)) {
  train[i, 'grass_weight'] <- weighting[train[i, 'surface'], 'Grass']
  train[i, 'clay_weight'] <- weighting[train[i, 'surface'], 'Clay']
  train[i, 'hard_weight'] <- weighting[train[i, 'surface'], 'Hard']
}

btm_grass <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(grass_weight), data = train)
btm_clay <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(clay_weight), data = train)
btm_hard <- BTm(outcome = 1, factor(winner_name, levels=players), factor(loser_name, levels=players), weights = abs(hard_weight), data = train)

# Create a dataframe with the Bradley-Terry estimates for each surface
btm_grass_df <- as.data.frame(BTabilities(btm_grass))
btm_clay_df <- as.data.frame(BTabilities(btm_clay))
btm_hard_df <- as.data.frame(BTabilities(btm_hard))

coeff_frames <- list('Grass' = btm_grass_df, 'Clay' = btm_clay_df,
                     'Hard' = btm_hard_df)

draws <- test[, c("winner_name", "loser_name", "surface")]
draws$pred <- NA

# predict the test set using a different model based on the surface
for (i in 1:nrow(draws)) {
  player1 <- draws$winner_name[i]
  player2 <- draws$loser_name[i]
  draw_surface <- draws$surface[i]

  pred <- predict(player1, player2, as.data.frame(coeff_frames[draw_surface]))
  draws$pred[i] <- pred[2]
}

score_predictions(draws)
```

This model performs better.
