# Recency Weighting

We want to weight recent matches more heavily than older matches.

Define

$$
W(t) = \min(f, f^t)
$$

where $f\in (0,1]$ is a weighting factor, and $t$ is the time since the last match. The min operation stops very recent matches being weighted far too high.

```{r}
library(tidyr)
library(dplyr)

time_weighting <- function(t, f) {
  return(min(f^t, f))
}

x <- seq(0,10,by=0.1)
plot(x, sapply(x, function(t) time_weighting(t,0.8)), type='l',
     ylab='weighting', xlab='time')
```

Let's import the data

```{r}
countsToBinomial<-function(xtab){
        ##assumes square
        players <- rownames(xtab)
        comb <- t(combn(nrow(xtab), 2))
        won <- xtab[comb]
        lost <- t(xtab)[comb]
        res <- !(won == 0 & lost == 0)
        player1 <- factor(players[comb[,1]], levels = players)[res]
        player2 <- factor(players[comb[,2]], levels = players)[res]
        return(data.frame(player1, player2, win1 = won[res], win2 = lost[res]))
}
```

```{r}
# Define the file path
file_paths <- c("../data/wta_matches_2022.csv", "../data/wta_matches_2021.csv", 
                "../data/wta_matches_2020.csv",
                "../data/wta_matches_2019.csv", "../data/wta_matches_2018.csv",
                "../data/wta_matches_2017.csv", "../data/wta_matches_2016.csv",
                "../data/wta_matches_2015.csv", "../data/wta_matches_2014.csv",
                "../data/wta_matches_2013.csv")

df1 <- read.csv(file_paths[1])

for (i in 2:(length(file_paths)-1)) {
  df1 <- rbind(df1, read.csv(file_paths[i]))
}

df1 <- df1[,c("winner_name","loser_name","tourney_date")]
```

Now remove players with fewer than 5 matches

```{r}
# Remove players in s that appear in fewer than 5 matches
players <- unique(c(df1$winner_name, df1$loser_name))
player_freqs <- setNames(rep(0, length(players)), players)
for (j in 1:length(players)) {
  for (k in 1:length(df1$winner_name)) {
    if (df1$winner_name[k] == players[j]) {
      player_freqs[j] <- player_freqs[j] + 1
    } else if (df1$loser_name[k] == players[j]) {
      player_freqs[j] <- player_freqs[j] + 1
    }
  }
}
```

```{r}
# filter the players to only have players with more than 5 matches
players <- names(player_freqs[player_freqs >= 20])
df1 <- df1[df1$winner_name %in% players & df1$loser_name %in% players, ]
df1
```

Now convert the tourney date into time since tournament in years.

```{r}
tourney_year = sapply(df1$tourney_date, function(x) {as.numeric(substr(x, 1, 4))})
time <- 2023-tourney_year
df1$time <- time
```

Now make the table to send to countsToBinomial(). We need to weight each match by it's recency weighting.

```{r}
mat <- matrix(0, nrow = length(players), ncol = length(players))
colnames(mat) <- players
rownames(mat) <- players

recency_weighting = 0.5

for (i in 1:nrow(df1)) {
  row <- df1[i,]
  mat[row$winner_name,row$loser_name] <- mat[row$winner_name,row$loser_name] + 1
}
```

Now countsToBinomial()

```{r}
tennis_data <- countsToBinomial(mat) 
head(tennis_data)
```

```{r}
# open the 2023 file and make predictions
matches_2023 <- read.csv("../data/atp_matches_2023.csv")
pred_df <- matches_2023

# make training data
matches_2023 <- matches_2023[matches_2023$winner_name %in% players & matches_2023$loser_name %in% players, ]
matches_2023 <- matches_2023[, c("winner_name", "loser_name")]
```

Make the model without time weighting for reference.

```{r}
library(BradleyTerry2)
players <- intersect(df1$winner_name, df1$loser_name)
btm <- BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    formula=~player, id='player', data=df1)
```

```{r}
BTabilities(btm)
```

```{r}
predict <- function(player1, player2) {
  lambda1 <- df_coeff[player1,1]
  lambda2 <- df_coeff[player2,1]

  pred <- exp(lambda1) / (exp(lambda1) + exp(lambda2))# Define the file path
file_paths <- c("../data/wta_matches_2022.csv", "../data/wta_matches_2021.csv", 
                "../data/wta_matches_2020.csv",
                "../data/wta_matches_2019.csv", "../data/wta_matches_2018.csv",
                "../data/wta_matches_2017.csv", "../data/wta_matches_2016.csv",
                "../data/wta_matches_2015.csv", "../data/wta_matches_2014.csv",
                "../data/wta_matches_2013.csv")

df1 <- read.csv(file_paths[1])

for (i in 2:(length(file_paths)-1)) {
  df1 <- rbind(df1, read.csv(file_paths[i]))
}

df1 <- df1[,c("winner_name","loser_name","tourney_date")]
  return(c(pred>0.5, pred))
}
```

```{r}
# open the 2023 file and make predictions
matches_2023 <- read.csv("../data/wta_matches_2023.csv")
matches_2023
# make training data
matches_2023 <- matches_2023[matches_2023$winner_name %in% players & matches_2023$loser_name %in% players, ]
matches_2023 <- matches_2023[, c("winner_name", "loser_name")]

```

```{r}
df_coeff <- as.data.frame(BTabilities(btm))


# make predictions
draws <- matches_2023[,c("winner_name","loser_name")]
draws <- draws[draws$winner_name %in% players & draws$loser_name %in% players, ]

for (i in 1:nrow(draws)) {
  player1 <- draws$winner_name[i]
  player2 <- draws$loser_name[i]
  pred <- predict(player1, player2)
  draws$pred[i] <- pred[2]
}
```

```{r}
library(rlist)
# find the accuracy
draws$correct <- draws$pred > 0.5
num <-nrow(filter(draws, draws$correct))
print(num / nrow(draws))
```

Now make the model with time weighting.

```{r}
recency_weighting <- 0.5
weights <- sapply(df1$time, function(x) time_weighting(x, recency_weighting))
```

```{r}
players <- intersect(df1$winner_name, df1$loser_name)
btm <- BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    weights = weights, id='player', data=df1)
```

```{r}
# open the 2023 file and make predictions
matches_2023 <- read.csv("../data/wta_matches_2023.csv")
# make training data
matches_2023 <- matches_2023[matches_2023$winner_name %in% players & matches_2023$loser_name %in% players, ]
matches_2023 <- matches_2023[, c("winner_name", "loser_name")]

```

```{r}
df_coeff <- as.data.frame(BTabilities(btm))


# make predictions
draws <- matches_2023[,c("winner_name","loser_name")]
draws <- draws[draws$winner_name %in% players & draws$loser_name %in% players, ]

for (i in 1:nrow(draws)) {
  player1 <- draws$winner_name[i]
  player2 <- draws$loser_name[i]
  pred <- predict(player1, player2)
  draws$pred[i] <- pred[2]
}
```

```{r}
library(rlist)
# find the accuracy
draws$correct <- draws$pred > 0.5
num <- nrow(filter(draws, draws$correct))
num / nrow(draws)
```

Much better :) Test all of them.

```{r}
for (recency_weighting in seq(0.1,1,0.1)) {
  cat(recency_weighting)
  weights <- sapply(df1$time, function(x) time_weighting(x, recency_weighting))
  btm <- BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    weights = weights, id='player', data=df1)
  
  df_coeff <- as.data.frame(BTabilities(btm))
  
  # make predictions
  draws <- matches_2023[,c("winner_name","loser_name")]
  draws <- draws[draws$winner_name %in% players & draws$loser_name %in% players, ]
  
  for (i in 1:nrow(draws)) {
    player1 <- draws$winner_name[i]
    player2 <- draws$loser_name[i]
    pred <- predict(player1, player2)
    draws$pred[i] <- pred[2]
  }
  
  draws$correct <- draws$pred > 0.5
  num <- nrow(filter(draws, draws$correct))
  cat(num / nrow(draws), "\n")
} 
```

So choose $f = 0.4$.

```{r}
standard_error <- function(x) sd(x) / sqrt(length(x))
```

```{r}
for (recency_weighting in seq(0.1,1,0.1)) {
  cat(recency_weighting)
  weights <- sapply(df1$time, function(x) time_weighting(x, recency_weighting))
  btm <- BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    weights = weights, id='player', data=df1)
  
  df_coeff <- as.data.frame(BTabilities(btm))
  
  # make predictions
  draws <- matches_2023[,c("winner_name","loser_name")]
  draws <- draws[draws$winner_name %in% players & draws$loser_name %in% players, ]
  
  for (i in 1:nrow(draws)) {
    player1 <- draws$winner_name[i]
    player2 <- draws$loser_name[i]
    pred <- predict(player1, player2)
    draws$pred[i] <- pred[2]
  }
  
  draws$correct <- draws$pred > 0.5
  num <- nrow(filter(draws, draws$correct))
  cat(" ", num / nrow(draws), " ",
      mean(draws[draws$pred > 0.5,]$pred), " ", standard_error(draws[draws$pred > 0.5,]$pred), " ",
      mean(log(draws[draws$pred > 0.5,]$pred)), " ", standard_error(log(draws[draws$pred > 0.5,]$pred)), "\n")
} 
```

| Time weighting | Accuracy | Average Probability | Average Log Probability |
|----------------|----------|---------------------|-------------------------|
| 0.1            | 0.640625 | 0.708 (0.004)       | -0.362 (0.006)          |
| 0.2            | 0.639375 | 0.700 (0.004)       | -0.372 (0.006)          |
| 0.3            | 0.63625  | 0.694 (0.004)       | -0.380 (0.005)          |
| 0.4            | 0.640625 | 0.688 (0.004)       | -0.389 (0.005)          |
| 0.5            | 0.63625  | 0.684 (0.004)       | -0.394 (0.005)          |
| 0.6            | 0.635625 | 0.679 (0.004)       | -0.401 (0.005)          |
| 0.7            | 0.6375   | 0.674 (0.003)       | -0.407 (0.005)          |
| 0.8            | 0.63375  | 0.673 (0.003)       | -0.410 (0.005)          |
| 0.9            | 0.631875 | 0.672 (0.003)       | -0.410 (0.005)          |
| 1              | 0.625625 | 0.674 (0.003)       | -0.407 (0.005)          |

```{r}
btm_w_1 = BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    weights = sapply(df1$time, function(x) time_weighting(x, 1)), id='player', data=df1)
as.data.frame(BTabilities(btm_w_1))[order(-BTabilities(btm_w_1)),]
```

```{r}
btm_w_01 = BTm(outcome=1, factor(winner_name, levels=players), factor(loser_name, levels=players), 
                    weights = sapply(df1$time, function(x) time_weighting(x, 0.1)), id='player', data=df1)
as.data.frame(BTabilities(btm_w_01))[order(-BTabilities(btm_w_01)),]
```
