---
title: "Model with surface"
author: "Grace Yan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing .csv files

First, we load the .csv files for all the WTA matches from 1968-present.

```{r, message=FALSE}
library(readr)
library(plyr)

myfiles <- list.files(path = "/home/cr23868/Documents/Tennis/R_code/tennis_wta-master", pattern = ".csv", full.names=TRUE)
dat_csv <- ldply(myfiles, read_csv)
View(dat_csv)
```

### Data wrangling

We extract the columns that we need. (We'll probably have to use EDA to justify the importance/unimportance of each variable.) These consist of: winner_hand, winner_ht, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon, w_SvGms, w_bpSaved, w_bpFaced, and their loser counterparts, and also surface.

Here, I will only include the winner_hand and winner_ht variables and their loser counterparts.

First, we extract players who have played at least 20 matches...

```{r, message=FALSE}
library(tidyr)
library(dplyr)

##### Extracting necessary columns and removing NAs
data <- data.frame(dat_csv[,c("winner_id", "winner_name", "winner_hand", "winner_ht", "w_ace", "w_df", "w_svpt", "w_1stIn", "w_1stWon",
                              "w_2ndWon", "w_SvGms", "w_bpSaved", "w_bpFaced", "loser_id", "loser_name", "loser_hand", "loser_ht",
                              "l_ace", "l_df", "l_svpt", "l_1stIn", "l_1stWon", "l_2ndWon", "l_SvGms", "l_bpSaved",
                              "l_bpFaced", "surface")])
data <- na.omit(data)
attach(data)

##### Extracting players who have won 20+ matches
all_players <- data.frame(Player=c(winner_name,loser_name))
count <- count(all_players, Player, name="Total_matches") # counts number of rows each unique name has in all_players, i.e. each unique player's number of total wins & losses
players <- filter(count, Total_matches >= 20) # extracts players who have played 20+ matches
```

Now we will extract the matches that we will use for our model based on two conditions: (i) only include matches where both the winner and loser's hand are known (i.e. 'winner_hand' and 'loser_hand' are not "U") and (ii) only include matches where both players have played 20+ matches in the dataset.

```{r}
# Indices on which players' hand is known
indices_w_handknown <- (winner_hand != "U") # TRUE for players whose hands are known, FALSE otherwise
indices_l_handknown <- (loser_hand != "U")
# Indices on which players have played 20+ matches
indices_w_handplayers <- winner_name %in% players$Player
indices_l_handplayers <- loser_name %in% players$Player

# Indices for rows in 'data' that have TRUE for all the indices above (i.e. matches where players' hands are known and they've played 20+ matches)
indices <- indices_w_handknown == TRUE & indices_l_handknown == TRUE & indices_w_handplayers == TRUE & indices_l_handplayers == TRUE

print( length(which(indices == TRUE)) ) # number of matches we will be using for the model
```

Using these indices, we will extract the rows we need from each variable.

```{r}
##### Extracting matches
matches <- data.frame(Winner=winner_name, Loser=loser_name)
matches <- matches[indices,]
# Convert characters into factors
matches$Winner <- factor(matches$Winner, levels=players$Player)
matches$Loser <- factor(matches$Loser, levels=players$Player)

are_levels_identical <- identical(levels(matches$Winner), levels(matches$Loser))
are_levels_identical

##### Extracting variables

# Extracting players' hand
W_Hand <- winner_hand[indices]
L_Hand <- loser_hand[indices]
# Convert characters into factors
W_Hand <- factor(W_Hand, levels = c('L', 'R'))
L_Hand <- factor(L_Hand, levels = c('L', 'R'))
# Make 'Hand' variable
Hand <- data.frame(W_Hand, L_Hand)

# Extracting players' heights
W_Height <- winner_ht[indices]
L_Height <- loser_ht[indices]
# Make 'Height' variable
Height <- data.frame(W_Height, L_Height)

# Extracting surface
Surface <- surface[indices]
Surface <- as.factor(Surface) # converting characters into factors
```

For player-specificity, we need to have an 'id' variable. The code below adds this column to the dataframe that we will use for the model.

```{r}
# Extracting players' ID
W_ID <- winner_id[indices] # only need W_ID since all the players we're considering have been winners

# Create an index for each player
unique_players <- unique(W_ID)
id <- match(W_ID, unique_players) # tells you which place in unique_players the elements in W_ID occupy.
```

So each time player i wins a match, the associated 'id' element will be the same. It is unique for each winner.

```{r}
# Combine everything into a new dataframe that will be used for modelling
data_for_model <- data.frame(id, matches, Hand, Height, Surface)
```

Fitting the Bradley Terry model with the hand, height and surface variables (all of them are player-specific):

```{r}
model_3vars <- BTm(1, Winner, Loser, ~ W_Hand[..] + L_Hand[..] + W_Height[..] + L_Height[..] + Surface[..], data = data_for_model)
model_3vars
summary(model_3vars)
```
