---
title: "Model with 2 covariates"
author: "Grace Yan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Importing .csv files

First, we load the .csv files for all the WTA matches from 1968-present.

```{r}
library(readr)
library(plyr)

myfiles <- list.files(path = "/home/cr23868/Documents/Tennis/R_code/tennis_wta-master", pattern = ".csv", full.names=TRUE)
dat_csv <- ldply(myfiles, read_csv)
View(dat_csv)
```

### Data wrangling

We extract the columns that we need. (We'll probably have to use EDA to justify the importance/unimportance of each variable.) These consist of: winner_hand, winner_ht, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon, w_SvGms, w_bpSaved, w_bpFaced, and their loser counterparts. (We'll add other variables like surface later - for now let's go with a simpler model as an experiment.)

Here, I will only include the winner_hand and winner_ht variables and their loser counterparts.

First, we extract players who have played at least 20 matches...

```{r}
library(tidyr)
library(dplyr)

##### Removing unnecessary columns and NAs
data <- data.frame(dat_csv[,c("winner_name", "winner_hand", "winner_ht", "w_ace", "w_df", "w_svpt", "w_1stIn", "w_1stWon",
                              "w_2ndWon", "w_SvGms", "w_bpSaved", "w_bpFaced", "loser_name", "loser_hand", "loser_ht",
                              "l_ace", "l_df", "l_svpt", "l_1stIn", "l_1stWon", "l_2ndWon", "l_SvGms", "l_bpSaved",
                              "l_bpFaced")])
data <- na.omit(data)
attach(data)

##### Extracting players who have won 30+ matches
all_players <- data.frame(Player=c(winner_name,loser_name))
count <- count(all_players, Player, name="Total_matches") # counts number of rows each unique name has in all_players, i.e. each unique player's number of total wins & losses
players <- filter(count, Total_matches >= 20) # extracts players who have played 20+ matches
```

Now we will have to remake the columns. Starting with 'winner_hand'...

```{r}
# Making a dataframe showing the hand of each player
hand_winners <- data.frame(Player=winner_name, Hand=winner_hand)
hand_losers <- data.frame(Player=loser_name, Hand=loser_hand)
hand <- bind_rows(hand_winners, hand_losers)
hand <- distinct(hand)

# Extracting players who played 30+ matches
indices1 <- hand$Player %in% players$Player
hand <- hand[indices1,]

# Getting rid of players whose hand is unknown
df_hand <- filter(hand, Hand != "U")
```

Then 'winner_ht'...

```{r}
height_winners <- data.frame(Player=winner_name, Height=winner_ht)
height_losers <- data.frame(Player=loser_name, Height=loser_ht)
height <- bind_rows(height_winners, height_losers)
height <- distinct(height)

# Making sure that the players in 'height' are only those already in our 'df_hand' dataframe
indices2 <- height$Player %in% df_hand$Player
df_height <- height[indices2,]

# Making sure that the players in 'df_hand' are the same as those in 'df_height'
indices0 <- df_hand$Player %in% df_height$Player
which(indices0 == FALSE) # none
```

We make a data frame called 'df_2vars' that consists of the players and the two variables 'Hand' and 'Height' that will be included in the model:

```{r}
# Bind columns
df_2vars <- left_join(df_hand, df_height, by="Player")

# Convert characters into factors
df_2vars$Player <- as.factor(df_2vars$Player)
df_2vars$Hand <- as.factor(df_2vars$Hand)
```

We also make a data frame called 'matches', consisting of all the matches that will be considered for the model.

```{r}
# Include only the matches where both players are in df_2vars
matches <- data.frame(Winner=winner_name, Loser=loser_name)

indices_w <- matches$Winner %in% df_2vars$Player
indices_l <- matches$Loser %in% df_2vars$Player
ind <- indices_w==TRUE & indices_l==TRUE

matches <- matches[ind,]

# Convert characters into factors
matches$Winner <- as.factor(matches$Winner)
matches$Loser <- as.factor(matches$Loser)
```

Now that we have extracted all the information we need from the dataset, we can make it into the following list:

```{r}
wta <- list(matches=matches, variables=df_2vars)
```

# Fitting the model using the BradleyTerry2 package

Using the `BTm` function in the BradleyTerry2 package, we fit the model:

```{r}
library(BradleyTerry2)
model_2vars <- BTm(1, Winner, Loser, ~ Hand[..] + Height[..], data = wta)
summary(model_2vars)
```
