---
title: "Version2"
author: "Grace Yan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing .csv files

First, we load the .csv files for all the WTA matches in 2013-2021. (Change path to the appropriate folder, which should contain .csv files of only the 2013-2021 matches.)

```{r, message=FALSE}
library(readr)
library(plyr)

myfiles <- list.files(path = "/home/cr23868/Documents/Tennis/R_code/tennis_wta-master", pattern = ".csv", full.names=TRUE)
dat_csv <- ldply(myfiles, read_csv)
```


## Creating the right format for the model

The process for reshaping the data is the following:

1. **Extract relevant columns**. We extract the columns that we will consider using: winner_id, winner_hand, winner_ht, w_ace, w_df, w_svpt, w_1stIn, w_1stWon, w_2ndWon, w_SvGms, w_bpSaved, w_bpFaced, their loser counterparts, and surface.
2. **Extract players who have played 20+ matches**.
3. **Reshape relevant variables**. Create variable columns only including the players who have played 20+ matches and whose hand is known (not 'U'). Convert characters to factors.
4. **Create dataframe for the model**. Create the final dataframe 'data_for_model' by binding the columns together in the form of 2 lists: 'contests' (containing winner's and loser's names and the Surface variable) and 'predictors' (the non match-specific variables).
5. **Bonus: create new variables**. We might want variables that do the following (inspired by the 'prev.wins' variables in chameleons dataset):

* Counts the number of previous contests each player has won
* Counts the number of aces etc. each player has achieved so far

### Step 1

Extracting the relevant columns...

```{r, message=FALSE}
library(tidyr)
library(dplyr)

##### Extracting necessary columns and removing NAs
data <- data.frame(dat_csv[,c("winner_id", "winner_name", "winner_hand", "winner_ht", "w_ace", "w_df", "w_svpt", "w_1stIn", "w_1stWon",
                              "w_2ndWon", "w_SvGms", "w_bpSaved", "w_bpFaced", "loser_id", "loser_name", "loser_hand", "loser_ht",
                              "l_ace", "l_df", "l_svpt", "l_1stIn", "l_1stWon", "l_2ndWon", "l_SvGms", "l_bpSaved",
                              "l_bpFaced", "surface")])
data <- na.omit(data)
attach(data)
```

### Step 2

We create 'players' which contains all the players who have played 20+ matches.

```{r}
# Extract all elements from winner_name and loser_name columns
all_players <- data.frame(Player=c(winner_name,loser_name))
# Count number of rows each unique player has in 'all_players' (i.e. each player's number of total wins & losses)
count <- count(all_players, Player, name="Total_matches")
# Only keep players who have played 20+ matches
players <- filter(count, Total_matches >= 20)

# Check that all the players in 'players' have won at least one match
which( !(players$Player%in%winner_name) )
```


### Step 4 

First we will extract the rows from 'winner_hand' and 'winner_ht' variables and their loser counterparts, based on two conditions: (i) only include matches where both the winner and loser's hand are known (i.e. 'winner_hand' and 'loser_hand' are not "U") and (ii) only include matches where both players have played 20+ matches in the dataset.

Starting with the 'hand' variable...

```{r}
# Make a dataframe showing the hand of each player
hand <- data.frame(Player=c(winner_name,loser_name), Hand=c(winner_hand,loser_hand))
hand <- distinct(hand) # eliminates duplicate rows
# Eliminate players whose hand is unknown
hand <- filter(hand, Hand != "U")
```

Then 'height' (a similar process to 'hand')...

```{r}
height <- data.frame(Player=c(winner_name,loser_name), Height=c(winner_ht,loser_ht))
height <- distinct(height)
```

Next, we make the 'predictors' dataframe. To do this, we first use `inner_join()` to combine all the columns we've made so far and naturally filter out players that don't fulfill conditions (i) and (ii). Then we add the ID column to 'predictors'.

```{r}
predictors <- players %>% inner_join(hand) %>% inner_join(height)
predictors = predictors[,-2]

# Extract ID and remove duplicated rows
id_df <- data.frame(ID=winner_id, Player=winner_name)
id_df <- distinct(id_df)
id_df <- id_df[order(id_df$Player),] # sorts ID from smallest to largest

# Join 'id_df' with 'predictors' and only keep players who are in both dataframes
predictors <- inner_join(id_df, predictors, by="Player")

# Convert characters to factors
predictors$Player <- factor(predictors$Player)
predictors$Hand <- factor(predictors$Hand)
predictors$ID <- factor(predictors$ID)
```

Next, we create the 'contest' dataframe where each row represents a match:

```{r}
# Create indices that allow us to extract matches where both the winner and loser must be in 'predictors$Player'
indices_w <- winner_name %in% predictors$Player
indices_l <- loser_name %in% predictors$Player
indices_wh <- (winner_hand != "U")
indices_lh <- (loser_hand != "U")
indices <- indices_w == TRUE & indices_l == TRUE & indices_wh == TRUE & indices_lh == TRUE

# Extract 'winner_name', 'loser_name' and 'surface' columns from 'data', and keep only the matches marked TRUE by indices
contests <- data.frame(Winner=factor(winner_name[indices]), Loser=factor(loser_name[indices]), Surface=factor(surface[indices]))
```


### Step 5

```{r}
# Convert the 'Player' column to rows
library(tibble)
predictors = column_to_rownames(predictors, var="Player")
```

```{r}
# Combine everything into a single list
data_for_model <- list(contests=contests, predictors=predictors)
View(data_for_model)
```

Note that the player identifier is 'ID' which contains the players' names. This allows R to apply player-specificity.

Fitting the Bradley Terry model with the hand, height and surface variables:

```{r}
model_Ha_He_Su <- BTm(player1 = Winner, player2 = Loser,
                      formula = ~ Hand[ID] + Height[ID] + Surface, id="ID",
                      data = data_for_model)
model_Ha_He_Su
summary(model_Ha_He_Su)
head(BTabilities(model_Ha_He_Su))

```

The 'surface' variable are NAs in the model when I format it this way. :( Alternatively, I tried to emulate the chameleons format by having Winner and Loser in two separate dataframes 'winner' and 'loser', and putting 'Surface' in both of them. Still gives NAs:

```{r}
winner <- contests[,c(1,3)]
loser <- contests[,c(2,3)]
# Rename
colnames(winner)[1] <- "Player"
colnames(loser)[1] <- "Player"

# No need to have ID column since Player will become our identifier in this new format
data_for_model2 <- list(winner=winner, loser=loser, predictors=predictors[,-1])
View(data_for_model)

model_Ha_He_Su2 <- BTm(player1 = winner, player2 = loser,
                      formula = ~ Hand[Player] + Height[Player] + Surface, id="Player",
                      data = data_for_model2)
model_Ha_He_Su2
summary(model_Ha_He_Su2)
head(BTabilities(model_Ha_He_Su2))
```

So both approaches ended in failure. And the previous version I did seemed tentative since every column's length was the number of matches - even the hand and height variables. I wonder why the prev.wins variables worked in the chameleons model but surface doesn't work here? They seem similar, except prev.wins are continuous I suppose?
