---
title: "Version3"
author: "Grace Yan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code from Version2 - just run without re-reading

First, we load the .csv files for all the WTA matches in 2013-2021. (Change path to the appropriate folder, which should contain .csv files of only the 2013-2021 matches.)

```{r, message=FALSE}
library(readr)
library(plyr)

myfiles <- list.files(path = "/home/cr23868/Documents/Tennis/R_code/tennis_wta-master", pattern = ".csv", full.names=TRUE)
dat_csv <- ldply(myfiles, read_csv)
```



```{r, message=FALSE}
library(tidyr)
library(dplyr)

##### Extracting necessary columns and removing NAs
data <- data.frame(dat_csv[,c("winner_id", "winner_name", "winner_hand", "winner_ht", "w_ace", "w_df", "w_svpt", "w_1stIn", "w_1stWon",
                              "w_2ndWon", "w_SvGms", "w_bpSaved", "w_bpFaced", "loser_id", "loser_name", "loser_hand", "loser_ht",
                              "l_ace", "l_df", "l_svpt", "l_1stIn", "l_1stWon", "l_2ndWon", "l_SvGms", "l_bpSaved",
                              "l_bpFaced", "surface")])
data <- na.omit(data)
attach(data)
```

Creating 'players' which contains all the players who have played 20+ matches.

```{r}
# Extract all elements from winner_name and loser_name columns
all_players <- data.frame(Player=c(winner_name,loser_name))
# Count number of rows each unique player has in 'all_players' (i.e. each player's number of total wins & losses)
count <- count(all_players, Player, name="Total_matches")
# Only keep players who have played 20+ matches
players <- filter(count, Total_matches >= 20)

# Check that all the players in 'players' have won at least one match
which( !(players$Player%in%winner_name) )
```

Starting with the 'hand' variable...

```{r}
# Make a dataframe showing the hand of each player
hand <- data.frame(Player=c(winner_name,loser_name), Hand=c(winner_hand,loser_hand))
hand <- distinct(hand) # eliminates duplicate rows
# Eliminate players whose hand is unknown
hand <- filter(hand, Hand != "U")
```

Then 'height' (a similar process to 'hand')...

```{r}
height <- data.frame(Player=c(winner_name,loser_name), Height=c(winner_ht,loser_ht))
height <- distinct(height)
```

Next, we make the 'predictors' dataframe:

```{r}
predictors <- players %>% inner_join(hand) %>% inner_join(height)
predictors = predictors[,-2]

# Extract ID and remove duplicated rows
id_df <- data.frame(ID=winner_id, Player=winner_name)
id_df <- distinct(id_df)
id_df <- id_df[order(id_df$Player),] # sorts ID from smallest to largest

# Join 'id_df' with 'predictors' and only keep players who are in both dataframes
predictors <- inner_join(id_df, predictors, by="Player")

# Convert characters to factors
predictors$Player <- factor(predictors$Player)
predictors$Hand <- factor(predictors$Hand)
predictors$ID <- factor(predictors$ID)
```

Then the 'matches' dataframe (this was called 'contests' in Version2 but I prefer the name 'matches'):

```{r}
# Create indices that allow us to extract matches where both the winner and loser must be in 'predictors$Player' and their hand is known
indices_w <- winner_name %in% predictors$Player
indices_l <- loser_name %in% predictors$Player
indices_wh <- (winner_hand != "U")
indices_lh <- (loser_hand != "U")
indices <- indices_w == TRUE & indices_l == TRUE & indices_wh == TRUE & indices_lh == TRUE

# Extract 'winner_name' and 'loser_name', and keeping only the matches marked TRUE by indices
matches <- data.frame( Winner=factor(winner_name[indices]), Loser=factor(loser_name[indices]) )
```


## Make continuous Surface variables

```{r}
# Extract Surface variable for the matches we're considering
Surface <- surface[indices]

# Creating 3 dataframes for matches played on clay, grass, hard court
ind_clay <- Surface == "Clay"
ind_grass <- Surface == "Grass"
ind_hard <- Surface =="Hard"
matches_clay <- matches[ind_clay,] # 5468 matches on clay
matches_grass <- matches[ind_grass,] # 2281 on grass
matches_hard <- matches[ind_hard,] # 13951 on hard court

# Initial values for the variables we're aiming to find
prev_wins_clay <- matrix(0, nrow=dim(matches_clay)[1], 2)
prev_wins_grass <-matrix(0, nrow=dim(matches_grass)[1], 2)
prev_wins_hard <- matrix(0, nrow=dim(matches_hard)[1], 2)

```

Below is the function for a 'for' loop that calculates the number of previous wins on a particular surface belonging to the winner and loser of each match.

```{r}

func_surface <- function(MATCHES, PREV_WINS) {
  
  players <- unique( c(MATCHES[,1], MATCHES[,2]) )
  scores <- data.frame(Player=players, Score=rep(0, length(players))) # Make vector of "scores" for each unique player.
  # 'scores' keeps track of their # previous wins on a surface.
  PREV_WINS <- matrix( 0, dim(MATCHES)[1], 2)
  
  for (n in 2:dim(PREV_WINS)[1]) {
    prev_winner <- which(scores$Player==MATCHES[n-1,1]) # gives row number of previous match winner in 'scores' df
    scores[prev_winner, 2] = scores[prev_winner, 2] + 1 # updates previous match winner's score
    
    curr_player1 <- which(scores$Player==MATCHES[n,1]) # gives row number of 1st player in upcoming match
    curr_player2 <- which(scores$Player==MATCHES[n,2]) # same for 2nd player
    PREV_WINS[n,] = c(scores[curr_player1, 2], scores[curr_player2, 2]) # gives current scores of players in upcomin match (unchanged if neither player were in previous match)
  }
  return(PREV_WINS)
}

```

Thus, we obtain the variables we need:

```{r}
func_clay <- func_surface(matches_clay, prev_wins_clay)
func_grass <- func_surface(matches_grass, prev_wins_grass)
func_hard <- func_surface(matches_hard, prev_wins_hard)

# Split 'matches' dataframe
winners <- data.frame(Player=matches[,1])
losers <- data.frame(Player=matches[,2])

# Add these variable columns to the 'winners' and 'losers' dataframes
# Hard court
winners$prev.wins.hard[which(ind_hard==TRUE)] <- func_hard[,1]
losers$prev.wins.hard[which(ind_hard==TRUE)] <- func_hard[,2]
# Clay
winners$prev.wins.clay[which(ind_clay==TRUE)] <- func_clay[,1]
losers$prev.wins.clay[which(ind_clay==TRUE)] <- func_clay[,2]
# Grass
winners$prev.wins.grass[which(ind_grass==TRUE)] <- func_grass[,1]
losers$prev.wins.grass[which(ind_grass==TRUE)] <- func_grass[,2]
```

Will try and fix the bugs with clay and grass ASAP.

## Fitting model

Make final dataframe:

```{r}
# Convert the 'Player' column to rows
library(tibble)
predictors = column_to_rownames(predictors, var="Player")
```

```{r}
# Combine everything into a single list
data_for_model <- list(winners=winners, losers=losers, predictors=predictors)
View(data_for_model)
```


```{r}
model_hardcourt <- BTm(player1=winners, player2=losers,
                  formula = ~ prev.wins.hard + Hand[Player] + Height[Player],
                  id="Player", data=data_for_model)

model_hardcourt
summary(model_hardcourt)
```

